Can We Crunch the Social Graph on the Cheap?

### 2011-01-23 doc/hotpar2011/evaluation.tex
multi-node case / delay sims

\subsection{Considering the 'multi-node case'}
	\todo{needs introduction, which will depend on what the system description section looks like}
	We assume that the QPI link to the network interface will be the minimum bandwidth and not the network. On the Nehalem, the QPI link to the network interface provides a bandwidth 12.8GB/s each way. This as much as 200Mref/s for 64 byte references. Given that half of the bandwidth will go towards satisfying remote requests, leaving half to making requests \todo{n-1/n req vs. sum-over-n(1/(n-1))}.  So, the cores must put out a bandwidth of 100Mref/s for full utilization of the link.  Assuming a reasonable network latency of $1-2\mu s$, by Little's Law, the number of outstanding references is $100Mref/s*1-2\mu s = 100-200$. This is the amount of memory concurrency that needs to be generated for the latency of memory requests to be hidden.
We consider two issues: first, can we support 100-200 misses?, and second, can the caches support enough contexts? \todo{Brandon: might want to be more explicit about the interplay between cache pressure, number of contexts, amount of references that can be generated by X number of contexts, etc}
	Can we support 100-200 misses? If the remote requests are reads that go to the global queue, then the number of outstanding references is limited to about 36. If the remote requests do not need to be stored in the global queue, then the line-fill buffers of the cores can support a total of $6 cores * 10 ref/core = 60 references$, still not quite enough. If, however, prefetches are messages--memory stores--requesting a piece of remote memory then many more outstanding references could be supported, either with a software data structure or a memory management device implemented in an FPGA.
\todo{talk about the simulated delay experiments? mention that there are still the same memory system limits like queues so results don't show what perf would look like beyond a certain point, but do show that by adding contexts we can achieve the same performance when latency goes up.}
	Can the caches support enough contexts? To start exploring this question, we did cache pressure experiments measuring the effect of working set size and number of software contexts on performance. Our experiments suggest that with up to about 8 contexts/core, we can achieve almost the same max bandwidth, and for one update per remote reference (expected for basic graph comps?), working set size/thread of up to 64KiB scales well. \todo{Brandon: need to be more specific.  Also, are there important highlights regarding how the plot scales as local updates increases}.  Threads do not take up much space in the cache, shown by XXX.

\todo{should we mention the lack of synchronization in the experiments and how intend to explore it in next step, given that it is part of our proposed runtime?}

\todo{order seems okay, but rework organization of subsections}

>>>>

\subsection{Expanding to Multiple Compute Nodes}
	\todo{better title?}\\
	\todo{needs introduction, which will depend on what the system description section looks like}\\
	We use the single processor experiments to give insight into concurrency requirements in a multi-node system that uses a network with relatively long latency.

	On the Nehalem, the QPI link to the network interface provides a bandwidth 12.8GB/s each way. This as much as 200Mref/s for 64-byte messages. For uniformly random distributed references across the entire shared memory space, we can expect half of this outgoing bandwidth will go towards satisfying remote requests, leaving half for making requests.  So, the cores can put out a bandwidth of 100Mref/s at full utilization of the link. Recent networks have suggested a max issue rate of 30M-90M messages per second is attainable \todo{cite mellanox press release and/or XT3/Eldorado} and this seems to be gradually increasing, so we use 100Mref/s as a good approximation to work with. Assuming a round-trip network latency of $1-2\mu s$, then by Little's Law, the number of outstanding references is $100Mref/s*1-2\mu s = 100-200$. \todo{Redo this calculation with proper numbers latency} This is the amount of memory concurrency that needs to exist for the latency of memory requests to be hidden when there is a rate of 100Mref/s.
We consider two issues: first, can we support 100-200 misses?, and second, can the caches support enough contexts? 
In order to keep the cores busy, there must be sufficient memory concurrency, and the number of contexts required to achieve this will vary. As there are more contexts in a core, there are more workings sets, including both data and coroutine stacks, that should fit in the cache to maintain performance.

	Can we support 100-200 misses? If the remote requests are reads that go to the global queue, then the number of outstanding references is limited to about 36. If the remote requests do not need to be stored in the global queue, then the line-fill buffers of the cores can support a total of $6 cores * 10 ref/core = 60 references$, still not quite enough. If, however, prefetches are messages--communicated with memory stores--requesting a piece of remote memory then many more outstanding references could be supported, either with a software data structure or a memory management device implemented in an FPGA.
	We performed some experiments with green threads on listwalk, where we simulate the network latency by not allowing a coroutine to continue after it does a prefetch-and-switch until some specified amount of time has passed.  As seen in Figure \todo{ref delay plot} it is still possible to achieve the maximum throughput: it just requires more contexts to supply enough concurrent references, as expected. This simulation is of course limited both by the fact that prefetched data is actually sitting in cache longer than it should (increasing the effective working set size) and that the memory system still caps performance, but it is a useful validation of adding software contexts to achieve the level of concurrency that is necessary.

	Can the caches support enough contexts? Having too many separate working sets that do not fit into cache will cause performance degradation and thrashing. To start exploring this question, we did cache pressure experiments measuring the effect of working set size and number of software contexts on performance. \todo{see carlson fig} Our experiments suggest that with up to about 8 contexts/core, we can achieve almost the same max bandwidth \todo{ as}, and for one update per remote reference \todo{expected for basic graph comps?}, working set size/thread of up to 64KiB scales well. \todo{Brandon: need to be more specific.  Also, are there important highlights regarding how the plot scales as local updates increases}.  Threads do not take up much space in the cache, shown by XXX.

\todo{should we mention the lack of synchronization in the experiments and how intend to explore it in next step, given that it is part of our proposed runtime?  Can we evaluate design choices qualitatively right now?}
